Containers and R---the Rockerverse and beyond

### 2020-01-15 manuscript.Rmd
some reformatting, add some links, some reordering, some labels

@@

**\pkg{stevedore}** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) 
<!-- from the CRAN description:
Work with containers over the Docker API. Rather than using system calls to interact with a docker client, using the API directly means that we can receive richer information from docker. The interface in the package is automatically generated using the 'OpenAPI' (a.k.a., 'swagger') specification, and all return values are checked in order to make them type stable.
-->

@@

**\pkg{harbor}** ([https://github.com/wch/harbor](https://github.com/wch/harbor)
<!-- from the github description
is an R package for controlling docker containers on local and remote hosts.
-->

@@

- [`dockerfiler`](https://github.com/ColinFay/dockerfiler/)
- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockr`](https://github.com/smaakage85/dockr), see also http://smaakage85.netlify.com/2019/12/21/dockr-easy-containerization-for-r/
- [`dockertest`](https://github.com/traitecoevo/dockertest/)
- [`renv` with Docker](https://rstudio.github.io/renv/articles/docker.html)
- https://github.com/gigantum/base-images

@@

## Deployment to cloud services

The cloud is the natural environment of containers, and becomes the go-to mechanism to expose applications written in R to users.

@@

@@

- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.
- Google Cloud Build
  - https://cloud.google.com/cloud-build/
  - Cloud Build is another continuous intergation service that helps move the workload of building the Docker images to an online service.  Cloud Build can be set up to build the Dockerfiles on each GitHub commit or release.  This means you do not need to build the Docker images locally, which can tie up resources since Docker images can be several GBs and take a long time to compile. Google Cloud Build works alongside Google Container Registry to allow you to build private and public Docker images, which allows you to build up your own dependency graphs for downstream applications.

@@

- Docker images for cloud services
  - A popular application of using Docker is to create the environments that can run in parallel for speeding up R code.  For example, [`googleComputeEngineR`](https://CRAN.R-project.org/package=googleComputeEngineR)'s `gce_vm_cluster()` function can create clusters of 2 or more virtual machines, running multi-CPU architectures.  Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running in a Docker container with the same R environments.   This is achieved through `googleComputeEngineR`'s integration with the R parralisation library [`library(future)` by Henrik Bengtsson]( https://CRAN.R-project.org/package=future).  Local R computation can be thrown up to a multi-CPU and VM environment to achieve parrell computation in a few lines of R code in your local session. - [some demonstrations are available here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html). 
- [Google Cloud Run](https://cloud.run) - This is a CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure.  This dispenses with the developer creating a cloud server to run the Docker image on, by abstracting away those servers to a more serverless configuration.  Cloud Run lets you run your code on top of a managed or your own Kubernetes cluster running Knative, and can accept any Docker image.  The service takes care of network ingress, scaling machines up and down to zero, authentication and authorisation, all features which are non-trivial for a developer to create on their own.  This can be used to scale up R code to millions of instances if they need to, with little or no changes to existing code. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses Cloud Run to create a scalable R plumber API. 
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g., start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.
- \pkg{drake}, https://docs.ropensci.org/drake/index.html?q=docker#with-docker

## Packaging programs for processing pipelines

@@

## Research Compendia

- https://github.com/benmarwick/rrtools
- \pkg{renv}, https://rstudio.github.io/renv/articles/docker.html

>>>>

**[Gigantum](http://gigantum.com/)** also its own image stack ([https://github.com/gigantum/base-images](https://github.com/gigantum/base-images)).
...

@@

**\pkg{harbor}** ([https://github.com/wch/harbor](https://github.com/wch/harbor) is not actively maintained anymore, but should be honorably mentioned as the first R package for managing Docker images and containers.
It uses the \pkg{sys} package to run system commands against the Docker CLI, both locally and through an ssh connection, and has convenience functions, e.g., for listing and removing containers/images and for accessing logs.
The output of container executions are converted to appropriate R types.
The Docker CLI, while evolving quickly and not with great concern for introducing breaking changes, basic functionality is unchanged for a long time so `harbor::docker_run(image = "hello-world")` works like a charm.

```{r harbor, eval=FALSE, include=FALSE}
remotes::install_github("wch/harbor")
harbor::docker_run(image = "hello-world")
```

**\pkg{stevedore}** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) 
<!-- from the CRAN description:
Work with containers over the Docker API. Rather than using system calls to interact with a docker client, using the API directly means that we can receive richer information from docker. The interface in the package is automatically generated using the 'OpenAPI' (a.k.a., 'swagger') specification, and all return values are checked in order to make them type stable.
-->
currently the most powerful Docker client in R.

@@

@@

**\pkg{dockerfiler}** ([https://github.com/ColinFay/dockerfiler/](https://github.com/ColinFay/dockerfiler/)) ...

**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) ...

**\pkg{dockr}** ([https://github.com/smaakage85/dockr](https://github.com/smaakage85/dockr)) ...
see/cite http://smaakage85.netlify.com/2019/12/21/dockr-easy-containerization-for-r/
<!-- from description on CRAN: Build a lightweight Docker r-base container image for your R package, in which the package itself is available. The Docker image seeks to mirror your R session as close as possible with respect to R specific dependencies. Both dependencies on CRAN R packages as well as local non-CRAN R packages will be included in the Docker container image.-->

- [`renv` with Docker](https://rstudio.github.io/renv/articles/docker.html)

@@

@@

The cloud is the natural environment of containers, and is the natural mechanism to deploy R server applications.

@@

- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.
- Google Cloud Build
  - https://cloud.google.com/cloud-build/
  - Cloud Build is another continuous intergation service that helps move the workload of building the Docker images to an online service.  Cloud Build can be set up to build the Dockerfiles on each GitHub commit or release.  This means you do not need to build the Docker images locally, which can tie up resources since Docker images can be several GBs and take a long time to compile. Google Cloud Build works alongside Google Container Registry to allow you to build private and public Docker images, which allows you to build up your own dependency graphs for downstream applications.
  - https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html

@@

- Docker images for cloud services
  - A popular application of using Docker is to create the environments that can run in parallel for speeding up R code.  For example, [`googleComputeEngineR`](https://CRAN.R-project.org/package=googleComputeEngineR)'s `gce_vm_cluster()` function can create clusters of 2 or more virtual machines, running multi-CPU architectures.  Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running in a Docker container with the same R environments.   This is achieved through `googleComputeEngineR`'s integration with the R parralisation library [`library(future)` by Henrik Bengtsson]( https://CRAN.R-project.org/package=future).  Local R computation can be thrown up to a multi-CPU and VM environment to achieve parrell computation in a few lines of R code in your local session. - [some demonstrations are available here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html). 
- [Google Cloud Run](https://cloud.run) - This is a CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure.  This dispenses with the developer creating a cloud server to run the Docker image on, by abstracting away those servers to a more serverless configuration.  Cloud Run lets you run your code on top of a managed or your own Kubernetes cluster running Knative, and can accept any Docker image.  The service takes care of network ingress, scaling machines up and down to zero, authentication and authorisation, all features which are non-trivial for a developer to create on their own.  This can be used to scale up R code to millions of instances if they need to, with little or no changes to existing code. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses Cloud Run to create a scalable R plumber API.
  - How does this relate to https://code.markedmondson.me/googleCloudRunner/index.html ?`
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g., start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.
- \pkg{drake}, https://docs.ropensci.org/drake/index.html?q=docker#with-docker

## Packaging programs for processing pipelines
\label{packaging}

@@

## Research Compendia
\label{compendia}

- researchcompendia.science
- https://github.com/benmarwick/rrtools
- \pkg{renv}, https://rstudio.github.io/renv/articles/docker.html
- what to copy in and what not
